---
title: "TCGA CNV of the genes in age-related losing interactions"
output: html_notebook
---

1.Setup the environment and load the tissue names
```{r}
library(dplyr)
library(data.table)
library(magrittr)
setwd("/data/user_data/wanght/CPM_methods/TCGA")
str_split(list.dirs("../Exps/")[2:27],pattern = "//",simplify = T)[,2] -> tissues

tissue_cancer<-fread("Cancer_tissue",header = T)
tissue_cancer<-tissue_cancer[tissue_cancer$Tissue %in% tissues,]
setkey(tissue_cancer,"TCGA_code")
cancers<-unique(tissue_cancer$TCGA_code)

TCGA_CNV<-fread("TCGA.PANCAN.sampleMap_Gistic2_CopyNumber_Gistic2_all_thresholded.by_genes.gz")
setkey(TCGA_CNV,"Sample")

survival_data<-fread("Survival_SupplementalTable_S1_20171025_xena_sp")
setkey(survival_data,"cancer type abbreviation")


```

##The survival status of samples with/wo CNV of genes involved in ARLI
2. Firstly, we focus on the CNV event that occurred in genes whom involved in the corresponding tissue's ARLI. For each cancer, we retrieved the matched tissue's ARLI, and divide the patients according to their tumor CNV events: one group with CNV in ARLI genes, the other not.
```{r}
library(stringr)

#Define the function to divide the patients
divide_cnv<-function(tcga){
  tissue<-tissue_cancer[tcga]$Tissue
  file_path<-paste0("../Result/", tissue,"/Result.RData")
  load(file_path)
  
  tissue_results_sig<-tissue_results[tissue_results$p<0.001 & tissue_results$tau>0,]
  LI_genes<-str_split(tissue_results_sig$edges,pattern = "_",simplify = T)[,1] %>% unique
  
  surv_data<-survival_data[tcga,c("sample","_PATIENT",
                                  "OS","OS.time","DSS","DSS.time","DFI","DFI.time","PFI","PFI.time")]
  surv_data<-surv_data[surv_data$sample %in% colnames(TCGA_CNV),]
  cnv_data<-TCGA_CNV[LI_genes] %>% select(surv_data$sample)
  apply(cnv_data,2,function(x){return(sum(x!=0,na.rm = T))})->surv_data$CNV_ARLI
  apply(TCGA_CNV%>% select(surv_data$sample),
        2,
        function(x){return(sum(x!=0,na.rm = T))})->surv_data$CNV_total
  
  return(surv_data)
}

TCGA_cnv_ARLI_res<-lapply(cancers, divide_cnv)
```

Then, let's perform the survival analysis for samples with more normalized ARLI genes CNV
```{r}
library(survival)
library(survminer)

tmpSurvdata<-TCGA_cnv_ARLI_res[[1]]
tmpSurvdata$risk<-ifelse(test = tmpSurvdata$CNV_ARLI/tmpSurvdata$CNV_total > 
                           median(tmpSurvdata$CNV_ARLI/tmpSurvdata$CNV_total,na.rm = T),
                         yes = "High", no ="Low")
survFit<-survfit(Surv(OS.time,OS==1)~risk,data=tmpSurvdata)
ggsurvplot(survFit,data=tmpSurvdata,conf.int = F,
           risk.table = F,pval=T)+labs(title=tissue_cancer[tissue_cancer$TCGA_code == cancers[1],]$Cancer_description %>%
                                         str_to_title())

tmp_plot<-function(num){
  tmpSurvdata<-TCGA_cnv_ARLI_res[[num]]
  tmpSurvdata$risk<-ifelse(test = tmpSurvdata$CNV_ARLI/tmpSurvdata$CNV_total > 
                             median(tmpSurvdata$CNV_ARLI/tmpSurvdata$CNV_total,na.rm = T),
                           yes = "High", no ="Low")
  survFit<-survfit(Surv(DSS.time,DSS==1)~risk,data=tmpSurvdata)
  ggsurvplot(survFit,data=tmpSurvdata,conf.int = F,
             risk.table = F,pval=T,palette = c("red","skyblue"))+
    labs(title=tissue_cancer[tissue_cancer$TCGA_code == cancers[num],]$Cancer_description %>% str_to_title())
}
```

