---
title: "Dosage sensitivity analysis of the genes in age-related losing interactions"
output: html_notebook
---

Collins et al ., present a beautiful work on [Cell](https://www.cell.com/cell/fulltext/S0092-8674(22)00788-7).They Discovered hundreds of rare CNV-disease associations across 54 disorders, and identified 3,635 highly dosage sensitive genes by scoring the orgnism's tolorence to thier CNV.

Due to the essence of the losing interaction is dosage imbalance, we wondered if the genes in age-related losing interactions would be more harmful when their dosage changed due to CNV events. 

We downloaded the data from [Zenodo](https://zenodo.org/record/6347673). Firstly, we will retrieve the dosage sensitive score of genes in age-related losing interactions to test whether they have a higher/lower CNV sensitive score (risk) than others. 

1.Setup the environment and load the tissue names
```{r}
setwd("/home/wanght/data/CPM_methods/Dosage")
library(data.table)
library(stringr)
library(magrittr)
str_split(list.dirs("../Exps/")[2:27],pattern = "//",simplify = T)[,2] -> tissues
tissues
```

## The Dosage scores
2.load the dosage table
```{r}
Dosage<-fread("Collins_rCNV_2022.dosage_sensitivity_scores.tsv.gz")
setkey(Dosage,"#gene")

#Define a function to retrieve the pHaplo and pTriplo from losing interact of tissues
#Input: tissue
#Output: the pHaplo/pTriplo scores of genes in OR not in losing interactions (LI)

dosage_score<-function(tissue){
  file_path<-paste0("../Result/", tissue,"/Result.RData")
  cat(tissue);cat(" ")
  load(file_path)
  tissue_results_sig<-tissue_results[tissue_results$p<0.001 & tissue_results$tau>0,]
  LI_genes<-str_split(tissue_results_sig$edges,pattern = "_",simplify = T)[,1] %>% unique
  nonLI_genes<-setdiff(str_split(tissue_results$edges,pattern = "_",simplify = T)[,1] %>% unique,
                       LI_genes)
  LI_genes_score<-Dosage[intersect(LI_genes,Dosage$`#gene`),];LI_genes_score$type<-"LI_genes"
  nonLI_genes_score<-Dosage[intersect(nonLI_genes,Dosage$`#gene`),];nonLI_genes_score$type<-"nonLI_genes"
  result<-rbind(LI_genes_score,nonLI_genes_score);result$tissue<-tissue
  return(result)
}

head(dosage_score(tissues[1]))
```

We then calculated the score among all tissues.
```{r}
Dosage_result<-lapply(tissues,
                      dosage_score)
names(Dosage_result)<-tissues
saveRDS(Dosage_result,file="Dosage_result_of_ARLI.RDS")
```



