---
title: "Dosage sensitivity analysis of the genes in age-related losing interactions"
output: html_notebook
---

Collins et al ., present a beautiful work on [Cell](https://www.cell.com/cell/fulltext/S0092-8674(22)00788-7).They Discovered hundreds of rare CNV-disease associations across 54 disorders, and identified 3,635 highly dosage sensitive genes by scoring the orgnism's tolorence to thier CNV.

Due to the essence of the losing interaction is dosage imbalance, we wondered if the genes in age-related losing interactions would be more harmful when their dosage changed due to CNV events. 

We downloaded the data from [Zenodo](https://zenodo.org/record/6347673). Firstly, we will retrieve the dosage sensitive score of genes in age-related losing interactions to test whether they have a higher/lower CNV sensitive score (risk) than others. 

1.Setup the environment and load the tissue names
```{r}
setwd("/home/wanght/data/CPM_methods/Dosage")
library(data.table)
library(stringr)
library(magrittr)
str_split(list.dirs("../Exps/")[2:27],pattern = "//",simplify = T)[,2] -> tissues
tissues
```

## The Dosage scores
2.load the dosage table
```{r}
Dosage<-fread("Collins_rCNV_2022.dosage_sensitivity_scores.tsv.gz")
setkey(Dosage,"#gene")

#Define a function to retrieve the pHaplo and pTriplo from losing interact of tissues
#Input: tissue
#Output: the pHaplo/pTriplo scores of genes in OR not in losing interactions (LI)

dosage_score<-function(tissue){
  file_path<-paste0("../Result/", tissue,"/Result.RData")
  cat(tissue);cat(" ")
  load(file_path)
  tissue_results_sig<-tissue_results[tissue_results$p<0.001 & tissue_results$tau>0,]
  LI_genes<-str_split(tissue_results_sig$edges,pattern = "_",simplify = T)[,1] %>% unique
  nonLI_genes<-setdiff(str_split(tissue_results$edges,pattern = "_",simplify = T)[,1] %>% unique,
                       LI_genes)
  LI_genes_score<-Dosage[intersect(LI_genes,Dosage$`#gene`),];LI_genes_score$type<-"LI_genes"
  nonLI_genes_score<-Dosage[intersect(nonLI_genes,Dosage$`#gene`),];nonLI_genes_score$type<-"nonLI_genes"
  result<-rbind(LI_genes_score,nonLI_genes_score);result$tissue<-tissue
  return(result)
}

head(dosage_score(tissues[1]))
```

We then calculated the score among all tissues.
```{r}
Dosage_result<-lapply(tissues,
                      dosage_score)
names(Dosage_result)<-tissues
saveRDS(Dosage_result,file="Dosage_result_of_ARLI.RDS")
```

For each tissue, we can test whether the genes in ARLI will have higher risk score.
```{r}
library(ggpubr)
library(gridExtra)
N<-1 # The Nth tissue
cat(tissues[N])

df<-Dosage_result[[N]]

p_hap <- ggboxplot(df, x = "type", y = "pHaplo",
                color = "type", palette =c("#FA0000", "#6666FF"),title = tissues[N])+
  stat_compare_means(comparisons = list(c("LI_genes","nonLI_genes")))+
  theme(legend.position = "none")
p_tri <- ggboxplot(df, x = "type", y = "pTriplo",
                color = "type", palette =c("#FA0000", "#6666FF"))+
  stat_compare_means(comparisons = list(c("LI_genes","nonLI_genes")))+
  theme(legend.position = "none")
grid.arrange(p_hap,p_tri,nrow=1)

#g<-arrangeGrob(p_hap,p_tri,nrow = 1)
#ggsave("Adipose.pdf",g)
```

```{r}
#OK, now we made them into a function
plotDosage<-function(N){
  df<-Dosage_result[[N]]

  p_hap <- ggboxplot(df, x = "type", y = "pHaplo",
                     color = "type", palette =c("#FA0000", "#6666FF"),title = tissues[N])+
           stat_compare_means(comparisons = list(c("LI_genes","nonLI_genes")))+
           theme(legend.position = "none")
  p_tri <- ggboxplot(df, x = "type", y = "pTriplo",
                color = "type", palette =c("#FA0000", "#6666FF"))+
            stat_compare_means(comparisons = list(c("LI_genes","nonLI_genes")))+
            theme(legend.position = "none")
  #grid.arrange(p_hap,p_tri,nrow=1)

  g<-arrangeGrob(p_hap,p_tri,nrow = 1)
  ggsave(paste0(tissues[N],".pdf"),g)
}
for(i in 1:length(tissues)){
    plotDosage(i)  
}
```

```{r}
D2<-do.call("rbind",Dosage_result)
ggboxplot(D2, x = "type", y = "pHaplo",
                     color = "type", palette =c("#FA0000", "#6666FF"))+ylim(c(0,1.5))+
           stat_compare_means(comparisons = list(c("LI_genes","nonLI_genes")))->p2
facet(p2,facet.by = "tissue",scales = "free_x")
```


Another plot method by ECDF cumulative curves
```{r}
ggecdf(D2,color="type",x="pTriplo",palette =c("#FA0000", "#6666FF"))+
  theme(axis.text = element_text(size=10))->p
facet(p,facet.by = "tissue",scales = "free_x")

```

##CNV-disease association
We have demonstrated that the Risk of dosage may be higher in ARLI genes. Then we wanna test whether the ARLI's dosage may be associated with some disorders
```{r}
library(dplyr)
Gene_asso<-read.table("Gene_associations_HPO.txt",sep = "\t",header = T)
phenotypes<-str_split(Gene_asso$Associated.HPOs,pattern = ";") %>% unlist %>% unique
phenos<-read.table("Collins_rCNV_2022.gene_association_sumstats/Phenotypes.txt",sep="\t",
                   header=T)

#Define the dosage association enrichment calculate functions
enrich_assoc<-function(pheno){
  cat(timestamp());cat("\n")
  dosage_gene<-Gene_asso[grep(Gene_asso$Associated.HPOs,pattern=pheno),] %>% 
    select(2:3) %>% apply(1,str_c,sep=";") %>% 
    str_split(pattern = ";") %>% unlist %>% unique
  dosage_gene<-dosage_gene[dosage_gene!="-"]
  
  res<-tibble(Tissue = tissues, OR=0, p=-1)
  for(i in 1:nrow(res)){
    #cat(timestamp())
    file_path<-paste0("../Result/", res[i,]$Tissue,"/Result.RData")
    cat(i);cat(" ")
    load(file_path)
    tissue_results_sig<-tissue_results[tissue_results$p<0.001 & tissue_results$tau>0,]
    LI_genes<-str_split(tissue_results_sig$edges,pattern = "_",simplify = T)[,1] %>% unique
    nonLI_genes<-setdiff(str_split(tissue_results$edges,pattern = "_",simplify = T)[,1] %>% unique,
                         LI_genes)
    tmpT<-tibble(genes = c(LI_genes, nonLI_genes),
                 inte = c(rep("LI",length(LI_genes)),
                          rep("nonLI",length(nonLI_genes))))
    tmpT$associat<-"non-Associated";tmpT[tmpT$genes %in% dosage_gene,]$associat<-"Associated"
    
    try({
      fit<-table(tmpT$inte,tmpT$associat) %>% fisher.test()
      fit$p.value->res[i,]$p; fit$estimate->res[i,]$OR},
      silent = T)
  }
  
  cat(timestamp())
  
  res$phenotype<-pheno
  return(res)
}

phenotypes<-phenotypes[phenotypes!="UNKNOWN"]
Enrich_res_gwas<-lapply(phenotypes, enrich_assoc)
save(Enrich_res_gwas,file="Enrich_res_rCNV_gwas.RData")
Enrich_res2<-do.call("rbind",Enrich_res_gwas)
```

```{r}
Enrich_res2<-do.call("rbind",Enrich_res_gwas)
Enrich_res2<-Enrich_res2[Enrich_res2$p>0,]
Enrich_res2[Enrich_res2$OR>100,]$OR<-100
rownames(phenos)<-phenos$HPO
phenos[phenotypes,]$Description
phenos[phenotypes,]$Description->pss
ggplot(Enrich_res_gwas[[6]],aes(x=Tissue,y=OR,fill=p<0.05))+
  scale_fill_manual(values=c("FALSE"="grey","TRUE"="red"))+
  geom_bar(stat = "identity") + coord_flip()+
  theme_classic()+ggtitle(pss[6])

ggplot(Enrich_res2,aes(x=Tissue,y=OR,fill=p<0.05))+
  geom_bar(stat = "identity") + 
  coord_flip()+theme_classic()+facet_wrap(~phenotype)
```
